\bfseries\LARGE\singlespacing{Transmission of respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-10-03 doc/paper/manuscript.tex
Updates from Overleaf

We analyzed 448~saliva samples over the study across classes. We detected 15~influenza~B~(Flu B), 15~rhinoviruses~(HRV), 14~adenovirus (AdV), 3~SARS-CoV-2~(CoV), 2~metapneumoviruses~(MPV), and 1~parainfluenza~virus~(PIV) (\Cref{fig:molecular-descriptives}a). There was an equal number of positive samples with and without air cleaners (\Cref{fig:molecular-descriptives}b). Taking into account the number of samples and the number of suceptible students, the positivity rate remained comparable between study conditions (adjusted relative risk 0.93, 95\%-Credible interval [CrI] 0.49 to 1.61), and was not sensitivity to infection to testing delays (Figure~\zref{fig:mol-estimation-results-sensitivity} in \supp). The distribution of positive samples varied between classes (\Cref{fig:molecular-descriptives}c). For example, in the first three study weeks all except for one sample were positive for AdV in class~A, while the vast majority of positive Flu~B samples were found in class~B. The virus-specific variations between classes are suggestive of transmission occurring within classes. Possible transmission chains within classes are shown in \Cref{fig:molecular-network}.

@@

We detected fewer respiratory viruses in bioaerosol samples compared to our previous study\cite{Banholzer2023PLoSMed} (1~sample of adenovirus and 1~sample of rhinovirus in this study vs 10~samples of SARS-CoV-2 in the previous study), especially in relation to the number of positive saliva samples (50~samples in this study vs 21~in the previous study). Similarly, the number of viruses detected on the filters of the air cleaners was also lower in this study (4~positive samples in 160~swabs vs 8~positive samples in 80~swabs in the previous study). Importantly, there was a clear shift in the distribution of respiratory viruses in the current study compared to the previous study in 2022 during the SARS-CoV-2 Omicron variant. Whereas in 2022 we detected almost exclusively SARS-CoV-2 in human saliva, the most common respiratory viruses in 2023 were adenovirus, influenza~B, and rhinoviruses (only three SARS-CoV-2-positive samples). The shift in the pattern of respiratory viruses has been observed in other studies\cite{Nygaard2023Lancet,Sauteur2022EuroSurv}. xxxxx

@@

Molecular transmission networks based on targeted sequencing or whole genome sequencing documented transmission in the classrooms. xxxx (influenza b, adeno, etc.). The low rate of positive bioaerosol samples also means that it is unclear to what extent students airborne transmission occurred in classrooms. It may be possible that students had rather little exposure to respiratory viruses in schools and acquired their infections elsewhere. However, the distribution of saliva samples between classes shows remarkable differences. Adenovirus was spreading entirely in class~A in the first three study weeks and only two infections with influenza~B occurred over the study. By contrast, influenza~B was spreading throughout the study in class~B and infections with adenovirus were only detected in the last study week. Adenovirus infections often tend to be mild\cite{Kunz2010CIDR} and less associated with coughing than influenza\cite{Ma2018RMV}, which is in line with the comparably lower frequency of detected coughs in class~B. Altogether, the class-specific patterns in the spread of respiratory viruses indicate that transmission of respiratory infections may have occurred at least to some extent within the classrooms.

>>>>

We analyzed 448~saliva samples over the study across classes. We detected 15~influenza~B~(Flu B), 15~rhinoviruses~(HRV), 14~adenovirus (AdV), 3~SARS-CoV-2~(CoV), 2~metapneumoviruses~(MPV), and 1~parainfluenza~virus~(PIV) (\Cref{fig:molecular-descriptives}a). There was an equal number of positive samples with and without air cleaners (\Cref{fig:molecular-descriptives}b). Taking into account the number of samples and the number of suceptible students, the positivity rate remained comparable between study conditions (adjusted relative risk 0.93, 95\%-Credible interval [CrI] 0.49 to 1.61), and was not sensitivity to infection to testing delays (Figure~\zref{fig:mol-estimation-results-sensitivity} in \supp). The distribution of positive samples varied between classes (\Cref{fig:molecular-descriptives}c). For example, in the first three study weeks all except for one sample were positive for AdV in class~A, while the vast majority of positive Flu~B samples were found in class~B. The virus-specific variations between classes are suggestive of transmission occurring within classes. Accordingly, we have manually labeled possible transmission chains within classes as shown in \Cref{fig:molecular-network}.

@@

We detected fewer respiratory viruses in bioaerosol samples compared to our previous study\cite{Banholzer2023PLoSMed} (1~sample of adenovirus and 1~sample of rhinovirus in this study vs 10~samples of SARS-CoV-2 in the previous study), especially in relation to the number of positive saliva samples (50~samples in this study vs 21~in the previous study). Similarly, the number of viruses detected on the filters of the air cleaners was also lower in this study (4~positive samples in 160~swabs vs 8~positive samples in 80~swabs in the previous study). Importantly, there was a clear shift in the distribution of respiratory viruses in the current study compared to the previous study in 2022 during the SARS-CoV-2 Omicron variant. Whereas in 2022 we detected almost exclusively SARS-CoV-2 in human saliva, the most common respiratory viruses in 2023 were adenovirus, influenza~B, and rhinoviruses (only three SARS-CoV-2-positive samples). The shift in the pattern of respiratory viruses has been observed in other studies\cite{Nygaard2023Lancet,Sauteur2022EuroSurv}.

@@

The low rate of positive bioaerosol samples also means that it is unclear to what extent students airborne transmission occurred in classrooms. It may be possible that students had rather little exposure to respiratory viruses in schools and acquired their infections elsewhere. However, the distribution of saliva samples between classes shows remarkable differences. For example, adenovirus was spreading entirely in class~A in the first three study weeks and only two infections with influenza~B occurred over the study. By contrast, influenza~B was spreading throughout the study in class~B and infections with adenovirus were only detected in the last study week. Adenovirus infections often tend to be mild\cite{Kunz2010CIDR} and less associated with coughing than influenza\cite{Ma2018RMV}, which is in line with the comparably lower frequency of detected coughs in class~B. Altogether, the class-specific patterns in the spread of respiratory viruses indicate that transmission of respiratory infections may have occurred at least to some extent within the classrooms.
